\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{hyperref}
\usepackage{amsmath, amssymb}
\usepackage{parskip}

% ─── Title Block ──────────────────────────────────────────────────────────────
\title{%
  \textbf{Interpretable N-Body Physics with GKANs}\\[0.4em]
  \large CPSC 440 Machine Learning --- Project Proposal
}
\author{
  Hanson Sun \quad Eitan Klinger \quad Arthur Guo\\
  \small\texttt{\{jsun33, ekling01, aguo07\}@cs.ubc.ca}
}

\begin{document}
\maketitle
\thispagestyle{empty}

% ─── 1. Problem Statement ─────────────────────────────────────────────────────
\section*{1. Problem Statement}

% What specific problem are you tackling?
% Introduce any non-ML background that a CS/stat/ECE grad student might not know.
% You do NOT need to introduce ML background covered in the course.

% [Describe the problem you are focusing on. Be precise: What is the input? What is the output or goal? Why is this problem interesting or important? Introduce any domain-specific background required to understand the setting.]

In their 2020 paper, ``\textit{Discovering Symbolic Models from Deep Learning with Inductive Biases}'',
Cranmer et al.\cite{cranmer2020discoveringsymbolicmodelsdeep} introduced a deep learning method for extracting relationships from simulated physics
interactions. Their work involved training a graph neural network (GNN) on the $n$-body problem then
performing symbolic regression to recover a generalizable analytical form for Newton's law of universal gravity.

One limitation of Cranmer et al.'s\cite{cranmer2020discoveringsymbolicmodelsdeep} approach is that the GNN is not directly interpretable. The additional symbolic regression step is computationally expensive, and the quality of the obtained analytical form
depends on whether the deep latent representation is semantically meaningful, as the learned parameters
may not correspond to real physical quantities like mass and velocity.

In 2024, Liu et al. demonstrated the potential of the Kolmogorov-Arnold network (KAN) to model physical
relationships in ``\textit{KAN}'' \cite{liu2025kankolmogorovarnoldnetworks} and ``\textit{KAN 2.0}'' \cite{liu2024kan20kolmogorovarnoldnetworks}. A KAN leverages the Kolmogorov-Arnold
representation theorem, which states that any multivariate function can be represented as a composition of univariate functions. While a traditional MLP has a fixed multivariate activation function at each node,
a KAN learns a univariate activation function for each edge \cite{liu2025kankolmogorovarnoldnetworks}. Liu et al.\cite{liu2025kankolmogorovarnoldnetworks}\cite{liu2024kan20kolmogorovarnoldnetworks} showed that KANs can outperform
MLPs in various cases and discussed the benefits of the function composition model in terms of interpretability.

Our goal is to explore the potential of a hybrid Kolmogorov-Arnold graph neural network (GKAN) architecture,
which replaces the multi-layer perceptron (MLP) layers in a GNN with KANs. We aim to prove this concept by training a GKAN on the
$n$-body problem following a similar approach as Cranmer et al.\cite{cranmer2020discoveringsymbolicmodelsdeep}, then evaluating the hybrid model against the
original GNN approach. We hypothesize that the GKAN can achieve comparable accuracy with the additional benefit
of being directly interpretable without requiring a symbolic regression step. We hope that this work contributes
to the ongoing research surrounding interpretability of deep models.


% ─── 2. Proposed Approach ────────────────────────────────────────────────────
\section*{2. Proposed Approach}

% What do you plan to do?
% Which methods will you apply, extend, or implement?
% What data will you use? How will you obtain / preprocess it?

[Describe your plan concretely. For example:
\begin{itemize}
  \item \textbf{Methods:} Which ML techniques will you use or build upon?
        (e.g., logistic regression, kernel methods, deep nets, graphical models)
  \item \textbf{Dataset:} What data will you work with, and how will you
        acquire / preprocess it?
  \item \textbf{Experimental setup:} How will you evaluate your approach?
        What baselines will you compare against?
\end{itemize}]

We will be using deep learning methods---specifically, we will focus on GNNs, KANs, and GKANs. We will
train using the $n$-body problem. The data will be generated via simulation using a numerically stable
ordinary differential equation (ODE) integrator such as leapfrog. We will evaluate our approach using quantitative metrics
such as training/validation loss, as well as qualitative metrics such as interpretability, and we will
compare our model against the baseline set by Cranmer et al.\cite{cranmer2020discoveringsymbolicmodelsdeep}.

In general, we aim to follow the baseline set by Cranmer et al.\cite{cranmer2020discoveringsymbolicmodelsdeep} quite closely, as we want to evaluate
whether the transition to KAN layers is a meaningful improvement compared to MLP-based approaches. As we
continue to develop and iterate on the approach, we may explore other physics problems, or add additional steps
such as pre-processing.


% ─── 3. Expected Contribution ────────────────────────────────────────────────
\section*{3. Expected Contribution}

% What will doing this project add to the world?
% Choose one of the project types below and explain your specific contribution.
%
% Common project types (pick one or mix):
%   Application bake-off | New application | Scaling up | Improving performance
%   Generalization to new setting | Survey/perspective | Coding project
%   Reproducibility report | Theory

%[State clearly what the contribution of your project will be. Examples:
%\begin{itemize}
%  \item Showing that method X outperforms Y on application Z, with an explanation of why.
%  \item A new algorithm that scales method X to datasets of size $n \gg 10^6$.
%  \item A structured survey of topic X with original insights on future directions.
% \item Negative result: a careful argument for why approach X does not work in setting Y, and why.
% \end{itemize}]

Our project is primarily an ``application bake-off'' (GKANs vs. GNNs), though it incorporates elements of
a few other project types.

Ideally, we will show that GKANs outperform GNNs with symbolic regression as a method of recovering
physical relationships, both in terms of accuracy and interpretability. We expect this to be the case
because the Kolmogorov-Arnold representation theorem framing allows us to choose our inductive biases
in a more physics-informed way, and the composition of univariate functions learned by the KAN layers can be directly visualized.


% ─── 4. Team \& Responsibilities (delete if solo) ────────────────────────────
\section*{4. Team \& Division of Responsibilities}

\begin{center}
\begin{tabular}{lll}
\hline
\textbf{Name} & \textbf{Student \#} & \textbf{Primary Responsibilities} \\
\hline
Author One   & 12345678 & [e.g., data preprocessing, experiments] \\
Eitan Klinger   & 41432253 & ...   \\
Author Three & 11223344 & [e.g., baseline comparisons, evaluation] \\
Arthur Guo & 16000796 & method implementation, experiments \\
\hline
\end{tabular}
\end{center}


% ─── 5. Timeline ─────────────────────────────────────────────────────────────
\section*{5. Timeline}

\begin{center}
\begin{tabular}{ll}
\hline
\textbf{Date}   & \textbf{Milestone} \\
\hline
Mar 29  & Proposal submitted \\
Apr 6   & [e.g., Data collected and preprocessed] \\
Apr 13  & [e.g., Baseline results obtained] \\
Apr 20  & [e.g., Main experiments complete] \\
Apr 27  & [e.g., Draft report finished] \\
[Final] & Report submitted \\
\hline
\end{tabular}
\end{center}


% ─── References ──────────────────────────────────────────────────────────────
\bibliographystyle{plain}
\bibliography{references}   % uncomment and point to your .bib file

% Manual references example (remove once you switch to BibTeX):

\end{document}
